WSClean: an implementation of a fast, generic wide-field imager

### 2014-02-25 articles/2014/Article-WSClean/wsclean.tex
Small updates

\subsection{Derivation of computing cost formulae}
Based on the expected cost terms described in Sec.~\ref{sec:time-complexity}, we derive analytical functions for the CASA and WSClean measurements using least-squares fitting. Several functions with different free parameters were tested, and formulae with minimum amount of parameters are selected that still follow the trend of the measurements and have reasonably small errors. Measurements are weighted with the inverse standard deviation. While it is normal to weight by inverse variance instead, we found that this results in too much weight on fast configurations, leading to functions that represent the general trend less well. The single outlying measurement for $N_\textrm{pix}=1792$ with CASA in Fig.~\ref{fig:timing-resolution} is removed.

@@

\subsection{Optimal snapshot duration}
Snapshot imaging, such as the recentring technique described in \S\ref{sec:snapshot-imaging}, can be implemented with either $w$-projection or $w$-stacking. Using the derived formulae, we can estimate the cost of making snapshots with both techniques. Snapshot imaging effectively removes the dependency on ZA from $w_{\max}$. Given the snapshot duration $\Delta \tau_\textrm{snapshot}$ and the total integration time $\Delta \tau_\textrm{total}$, the total time cost of inversion becomes
\begin{equation} \label{eq:snapshot-cost}
t_\textrm{snapshot}(\Delta \tau_\textrm{snapshot}) = \frac{\Delta \tau_\textrm{total}}{\Delta \tau_\textrm{snapshot}} t_\textrm{imaging}(w'_{\max},N'_\textrm{vis},N_\textrm{freq},N_\textrm{pix}),
\end{equation}
with
\begin{equation}
w'_{\max} = \frac{1}{\lambda}\max \Delta z \left(1.0 - \cos(\frac{1}{2}(\textrm{FOV} + \omega_E \tau_\textrm{snapshot})\right),
\end{equation}
$\omega_E$ the rotational speed of the Earth and $N'_\textrm{vis} = N_\textrm{vis}\frac{\Delta \tau_\textrm{total}}{\tau_\textrm{snapshot}}$. We have excluded the cost of phase shifting the visibilities and gridding. The cost for regridding with simple nearest neighbour or bilinair interpolation is indeed negligable, although more accurate interpolation (e.g. Lanczos interpolation) can be expensive. Also, our current implementation of the phase-changing program does take a non-negligable time, but this implementation is not optimized and can in theory be implemented in a preprocessing pipeline or on the fly during imaging.

@@

So far, we have corrected the (full-polarization) beam in image space, which is possible because of the homogenuity of the MWA tiles. This is cheap for our current snapshot time of 112s, but this might be unfeasible when it is required to calculate the beam on very short time scales. Accurate MWA beam models are still being developed. When beam updates are required often, calculating $a$-projection kernels or performing snapshot imaging also becomes more expensive, and it would be interesting to find out where the balance lies between these methods. When only a few thousand components need to be deconvolved, an approach with direct Fourier transforms is accurate and affordable. For wide-field arrays this can be combined with calibration of the direction dependence, e.g. with the SAGECAL \citep{sage-calibration-ii} or RTS peeling \citep{rts-mwa} calibration techniques. This is not possible for fields with diffuse emmission for which no model is known a-priory.

>>>>

\subsection{Derivation of computing cost formulae}
Based on the expected cost terms described in Sec.~\ref{sec:time-complexity}, we derive analytical functions for the CASA and WSClean measurements using least-squares fitting. Several functions with different free parameters were tested, and formulae with minimum amount of parameters are selected that still follow the trend of the measurements and have reasonably small errors. Measurements are weighted with the inverse standard deviation instead of the variance, because we found that the latter results in too much weight on fast configurations, leading to functions that represent the general trend less well. The single outlying measurement for $N_\textrm{pix}=1792$ with CASA in Fig.~\ref{fig:timing-resolution} is removed.

@@

\subsection{Optimal snapshot duration}
Snapshot imaging, such as the recentring technique described in \S\ref{sec:snapshot-imaging}, can be implemented with either $w$-projection or $w$-stacking. Using the derived formulae, we can estimate the cost of making snapshots with both techniques. Snapshot imaging effectively removes the dependency on ZA from $w_{\max}$. Given the snapshot duration $\Delta \tau_\textrm{snapshot}$ and the total integration time $\Delta \tau_\textrm{total}$, the total time cost of inversion becomes
\begin{equation} \label{eq:snapshot-cost}
t_\textrm{snapshot}(\Delta \tau_\textrm{snapshot}) = \frac{\Delta \tau_\textrm{total}}{\Delta \tau_\textrm{snapshot}} t_\textrm{imaging}(w'_{\max},N'_\textrm{vis},N_\textrm{freq},N_\textrm{pix}),
\end{equation}
with
\begin{equation}
w'_{\max} = \frac{1}{\lambda}\max \Delta z \left(1.0 - \cos(\frac{1}{2}(\textrm{FOV} + \omega_E \Delta \tau_\textrm{snapshot})\right),
\end{equation}
$\omega_E$ the rotational speed of the Earth and $N'_\textrm{vis} = N_\textrm{vis}\frac{\Delta \tau_\textrm{total}}{\tau_\textrm{snapshot}}$. We have excluded the cost of phase shifting the visibilities and gridding. The cost for regridding with simple nearest neighbour or bilinair interpolation is indeed negligable, although more accurate interpolation (e.g. Lanczos interpolation) can be expensive. Also, our current implementation of the phase-changing program does take a non-negligable time, but this implementation is not optimized and can in theory be implemented in a preprocessing pipeline or on the fly during imaging.

@@

So far, we have corrected the (full-polarization) beam in image space, which is possible because of the homogenuity of the MWA tiles. This is cheap for our current snapshot time of 112s, but this might be unfeasible when it is required to calculate the beam on very short time scales. Accurate MWA beam models are still being developed. When beam corrections are required on short time scales, calculating $a$-projection kernels or performing snapshot imaging also becomes more expensive, and it would be interesting to find out where the balance lies between these methods. When only a few thousand components need to be deconvolved, an approach with direct Fourier transforms is accurate and affordable. For wide-field arrays this can be combined with calibration of the direction dependence, e.g. with the SAGECAL \citep{sage-calibration-ii} or RTS peeling \citep{rts-mwa} calibration techniques. This is not possible for fields with diffuse emmission or faint point sources when no model is known a-priory.
